Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-19 shanghai.tex
Working on functoriality

The main result of this paper is a canonical isomorphism
\[
\CSiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.
  \Clifton{I have not proved $(f\circ g)^* = g^* \circ f^*$, yet.}
\end{proof}

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$.

@@

>>>>

The main result of this paper is a canonical isomorphism
\[
\CSiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about pullbacks of sheaves and maps of sheaves.
  \David{Is this explanation of $(f\circ g)^* = g^* \circ f^*$ sufficient?}
\end{proof}

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; the snake lemma then allows us to find $\CSiso{G}$.

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
We define a functor $\CS(G_{k'}) \to \CS(G)$ by.... \David{We can't define this by pullback
since $G_{k'} \to G$ goes in the wrong direction}
\begin{proposition} \label{prop:csbe}
The diagram
\[
\begin{tikzcd}
\CSiso{G_{k'}} \arrow[two heads]{r} \dar & \CSiso{G} \dar \\
\Hom(G(k'), \EEx) \arrow[two heads]{r} & \Hom(G(k), \EEx)
\end{tikzcd}
\]
commutes, where the vertical arrows are as in Theorem \ref{thm:snake}.
\end{proposition}

Now suppose that $G'$ is an abelian group scheme over $k'$, satisfying the conditions of Theorem \ref{thm:snake}.
There is a canonical isomorphism between $G'(k')$ and $(\Res_{k'/k}G')(k)$;
we now describe the corresponding functor from
$\CS(G')$ to $\CS(\Res_{k'/k}G')$.  The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
The component corresponding to the identity element yields a natural inclusion 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
of $k'$-schemes.  The definition of base change yields a map of $k$-schemes
\[
(\Res_{k'/k}G')_{k'} \to \Res_{k'/k}G'.
\]
Consecutive pullback along these maps yields a functor
\[
\CS(G') \to \CS(\Res_{k'/k}G').
\]

\begin{proposition}
  Then the following diagram commutes:
\[
\begin{tikzcd}
\CSiso{G} \rar{t_G} \dar & \Hom(G(k'), \EEx) \dar \\
\CSiso{\Res_{k'/k} G} \rar{t_{\Res_{k'/k} G}} & \Hom((\Res_{k'/k}G)(k), \EEx)
\end{tikzcd}
\]
\end{proposition}
